---
title: "STAT2006 FINAL ASSESMENT"
output: html_document
---

#### Group Members:
- **Leo Timmins** (*23559213*)
- **Noah Grieve** (*Number here*)
  
#### Notes:
- Project tracked with git
- Datasets found under `./data/`

#### Librarys
```{r}
#| warning: false
#| message: false # just to get rid of an error message im having with conflicts
library(tidyverse)
library(tidymodels)
library(leaps)
```

### Question 1.

#### Section 1a.
Read the dataset project_data1.csv into R and perform an 80/20 training–test split using a random
seed of 2006. Clearly show the training and test datasets created. (5 marks)
```{r}
# read the dataset ./data/project_data1.csv
project_data1 <- read_csv("./data/project_data1.csv")

# preform an 80/20 split with seed of 2006
set.seed(2006)

split_data1 <- initial_split(project_data1, prop = 0.8)
train_data1 <- training(split_data1)
test_data1 <- testing(split_data1)

# clearly show the training and test datasets
print(train_data1)
print(test_data1)
```

Using the training dataset, perform forward stepwise variable selection with the regsubsets()
function. Use all available predictors and set the maximum number of variables to 25. (5 marks)
```{r}
forward_stepwise <- regsubsets(
  y ~ .,                          # Y is predicted by the 25 X columns
  data = train_data1,             # uses the training dataset
  nvmax = 25,                     # set the maximum number of variables to 25
  method = "forward"              # forward stepwise variable selection
)

forward_summary <- summary(forward_stepwise)
```

Obtain the Cp values from the fitted forward stepwise model and produce a plot of Cp versus the
number of predictors 𝑝. Clearly highlight the model corresponding to the minimum Cp value on
the plot. (4 marks)
```{r}
cp_vals <- forward_summary$cp
predictors <- 1:length(cp_vals)

plot(
  predictors,                            # X data
  cp_vals,                               # Y data
  type = "b",                            # shows both points and lines
  xlab = "Number of Predictors",         # X label
  ylab = "Cp",                           # Y label
  main = "Cp vs. Number of Predictors"   # Title
)

min_cp <- which.min(cp_vals) # Finds lowest Cp value

# changes the point to red
points(
  predictors[min_cp],
  cp_vals[min_cp],
  col = "red",
)
```


#### Section 1b.
Using the training dataset obtained in Q1(a), perform the following analyses using 5-fold cross-validation
with a random seed of 2006.

Fit the following penalized regression models using cv.glmnet():

- Lasso regression (𝛼 = 1)
- Ridge regression (𝛼 = 0)
- Elastic-net regression (𝛼 = .5)

Use the same fold allocation (hint: see foldid argument) for all three models and produce the
corresponding cross-validation plots. (6 marks)
```{r}
```

Extract the lambda.1se value for each of the three fitted models and fit the final lasso, ridge, and
elastic-net models using the extracted tuning parameters. (6 marks)
```{r}
```

Fit a linear regression model using the predictors selected from the forward stepwise selection
procedure. (2 marks)
```{r}
```

Use the test dataset to obtain predictions for all four fitted models:
– Lasso
– Ridge
– Elastic-net
– Linear regression model selected using forward selection
Compute and present the RMSE values for all four models in a suitable table ordered from the
smallest RMSE to the largest RMSE. (5 marks)
```{r}
```

Based on the RMSE values, identify the best-performing model. (2 marks)
```{r}
```

### Question 2.
Read the dataset project_data2.csv into R and perform a 75/25 training–test split using a random
seed of 2006. (5 marks)
```{r}
```

Using the training dataset, fit elastic-net regression models using cv.glmnet() for the following
values of 𝛼: 𝛼 = 0.0, .1 , .2, ..., 1.0 

Store the fitted cross-validation objects using a tibble list-column approach. (5 marks)
```{r}
```

Extract the lambda.1se value corresponding to each value of 𝛼𝛼 and fit the associated elastic-net
model using glmnet(). Store the fitted models in tibble list-columns. (7 marks)
```{r}
```

Use the fitted models to predict the response values for the test dataset and compute the RMSE
for each fitted model using a tibble list-column workflow. Present the RMSE values in ascending
order and identify the optimal elastic-net model. (5 marks)
```{r}
```

For the optimal elastic-net model, produce a scatter plot of predicted versus actual response
values for the test dataset. Include the reference line 𝑦 = 𝑥 on the plot. (3 marks)
```{r}
```

### Question 3.
The dataset heart-disease.csv is available in the Datasets folder. In this question, your objective is to build
classification models for predicting the variable target.

The variables in the dataset are described below.

- age: age in years
- sex: sex (1 = male; 0 = female)
- cp: chest pain type
- trestbps: resting blood pressure (in mm Hg on admission to the hospital)
- chol: serum cholestoral in mg/dl
- fbs: fasting blood sugar > 120 mg/dl (1 = true; 0 = false)
- restecg: resting electrocardiographic results
- thalach: maximum heart rate achieved
- exang: exercise induced angina (1 = yes; 0 = no)
- oldpeak: ST depression induced by exercise relative to rest
- slope: the slope of the peak exercise ST segment
- ca: number of major vessels (0–3) coloured by fluoroscopy
- thal: 0 = normal; 1 = fixed defect; 2 = reversible defect
- target: 1 = presence of heart disease; 0 = absence of heart disease

For this question, you must use the tidyverse list-column approach with tibble, as demonstrated in the
lectures and computer laboratories. In particular, you should use functions such as:

- map()
- map2()
- map_dbl()
- crossing()

from the purrr and tidyr packages to store fitted models, predictions, confusion matrices, and performance
metrics.

Read the dataset heart-disease.csv into R and convert the variables target, sex, fbs, exang, ca, and
thal into factors. Display an appropriate summary of the dataset. (3 marks)
```{r}
```

Using a random seed of 2006, divide the data into an 80/20 training–test split using stratified
sampling based on target. Verify that the proportions of target = 0 and target = 1 are similar in the
original, training, and test datasets. (7 marks)
```{r}
```

Perform 5-fold cross-validation on the training dataset using stratified folds based on target.
Consider the following values of mtry for the random forest model
mtry = 2,3,4,5,6,7,8,9,10,11,12

Use the crossing() function to create a tibble containing all combinations of cross-validation folds and
mtry values. (5 marks)
```{r}
```

For each combination of cross-validation fold and mtry value:
– Create the corresponding training and validation datasets,
– Fit a random forest classification model using the training fold,
– Obtain predictions for the validation fold.

Store all fitted models and predictions using tibble list-columns. (5 marks)
```{r}
```

For each fitted random forest model, compute the following performance measures based on the
validation fold predictions:
– Accuracy
– Sensitivity
– Specificity

Then compute the average value of each metric across the five cross-validation folds for every mtry
value. Present the results in a suitable table ordered by decreasing accuracy. Identify the optimal mtry
value. (10 marks)
```{r}
```

Using the entire training dataset:
– Fit a logistic regression model,
– Fit a random forest model using the optimal mtry value selected from the cross-validation
procedure.

Evaluate both models on the test dataset and report the following performance metrics:
- Accuracy
- Sensitivity
- Specificity
Also compute the AUC values and produce ROC curves for both models. Based on the overall results,
comment on which model you would recommend for predicting heart disease. (10 marks)

```{r}
```


